Mining the human microbiome for clinical insight

### 2018-11-29 chap1.tex
Small edits to meta analysis intro

In Chapter 3, I perform a meta-analysis of 28 case-control gut microbiome studies across 10 diseases to synthesize findings across studies and identify generalizable associations.
Although the human gut microbiome has been extensively studied in many diseases, there is little consensus on which disease associations are consistent across patient cohorts.
In other fields like medicine or psychology, consensus is usually achieved through meta-analyses of published literature.
However, comparing published results across microbiome studies is not straightforward.
The field lacks standard data processing and analysis methods, making many reported results impossible or inappropriate to compare directly between studies.
For example, different studies often use incompatible bioinformatics or statistical methods: studies may compare bacteria at different taxonomic levels or use incomparable bioinformatics workflows, and they may identify significant associations through univariate statistical tests or machine learning models, all of which complicates the comparison of results across studies.
Additionally, studies led by clinicians often ask very different questions of the data than studies led by microbial ecologists, and so the reported results are not necessarily representative of the comprehensive information capable of being extracted from the full dataset.
This issue is especially pertinent to the gut microbiome field, where stool sample collection and open-source bioinformatics software suites have made microbiome research accessible to a broad range of scientists and clinicians alike.

In this chapter, I present a meta-analysis of gut microbiome studies which overcomes many of these challenges by reprocessing and reanalyzing data with standard methods.
Even though specific bacterial associations may vary across studies of the same disease, I do find consistent patterns of general microbiome shifts which are consistent.
These shifts suggest different associated approaches for developing microbiome-based therapeutics.
When looking across multiple diseases, we find a set of bacteria which are non-specifically associated with health and disease, suggesting a shared response to health and disease.
These results highlight the importance of contextualizing individual study results within existing work and also hints at the possibility of developing broadly beneficial targeted probiotic and antibiotic therapies.

>>>>

In Chapter 3, I present a meta-analysis of 28 case-control gut microbiome studies across 10 diseases to synthesize findings across studies and identify generalizable associations.
Although the human gut microbiome has been extensively studied in many diseases, there is little consensus on which disease associations are consistent across patient cohorts.
In other fields like medicine or psychology, consensus is usually achieved through meta-analyses of published literature.
However, comparing published results across microbiome studies is not straightforward.
The field lacks standard data processing and analysis methods, making many reported results impossible or inappropriate to compare directly between studies.
For example, different studies often use incompatible bioinformatics or statistical methods: they may compare bacteria at different taxonomic levels or use incompatible bioinformatics workflows, and they may identify significant associations through univariate statistical tests or machine learning models, results from which can not be compared.
Additionally, studies led by clinicians often ask very different questions of the data than studies led by microbial ecologists, and so the reported results are not necessarily representative of the potential information contained in the full dataset.
This issue is especially relevant in this field, where non-invasive sample collection and open-source bioinformatics software suites have made microbiome research accessible to a broad range of scientists and clinicians.

In this chapter, I perform a meta-analysis of gut microbiome studies which overcomes many of these challenges by reprocessing and reanalyzing data with standard methods.
Even though specific bacterial associations vary across studies of the same disease, I identify patterns of general microbiome shifts which are consistent  and which suggest different approaches for developing microbiome-based therapeutics.
When looking across multiple diseases, I find a set of bacteria which are non-specifically associated with health and disease, perhaps forming a shared or core response to health and disease.
These results highlight the importance of contextualizing results from individual studies within the existing body of work, and also hints at the possibility of developing broadly beneficial targeted probiotic and antibiotic therapies.
